\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}

### 2016-02-02 draft.tex
move GH puzzle text to end of intro

Consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size. One simple illustration from the realm of trade policy is the following: a legislator whose district has a large concentration of a particular industry does not take into account the impact of tariffs on the welfare of consumers in other districts, while the executive, whose constituency encompasses the whole country, will internalize these diffuse consumption effects. See Lohmann and O'Halloran (1994) for a detailed argument and Grossman and Helpman (2005) for a formal model that demonstrates the point for the legislature.\label{fn:ga_l_e}} For any trade-agreement tariff set by the executives, the lobby can calculate the effort level required to prevent the agreement from being ratified: it must shift the preferences of the median legislator just far enough so that he will choose the non-cooperative tariffs over the trade agreement tariffs. Because the lobby will pay this price so long as its benefit outweighs the cost, a trade war will occur unless the executive sets the trade-agreement tariff so that lobbying is not worthwhile.\footnote{As will be made clear below, non-ratification is equivalent to breaking a trade agreement once in place; likewise, the non-cooperative tariffs that result from failing to ratify the trade agreement are the same as the tariffs that would arise in trade war, so this model applies to both situations.}

>>>>

\vskip.2in
Consider first the case of political certainty, and assume that the executive is more pro-social than the legislature, as has been the case throughout the post-war period in the United States.\footnote{The standard assumption that the executive is less protectionist than the legislature is a special case of the finding that susceptibility to special interests generally declines with constituency size. One simple illustration from the realm of trade policy is the following: a legislator whose district has a large concentration of a particular industry does not take into account the impact of tariffs on the welfare of consumers in other districts, while the executive, whose constituency encompasses the whole country, will internalize these diffuse consumption effects. See Lohmann and O'Halloran (1994) for a detailed argument and Grossman and Helpman (2005) for a formal model that demonstrates the point for the legislature.\label{fn:ga_l_e}} For any trade-agreement tariff set by the executives, the lobby can calculate the effort level required to prevent the agreement from being ratified: it must shift the preferences of the median legislator just far enough so that he will choose the non-cooperative tariffs over the trade agreement tariffs. Because the lobby will pay this price so long as its benefit outweighs the cost, a trade war will occur unless the executive sets the trade-agreement tariff so that lobbying is not worthwhile.\footnote{As will be made clear below, non-ratification is equivalent to breaking a trade agreement once in place; likewise, the non-cooperative tariffs that result from failing to ratify the trade agreement are the same as the tariffs that would arise in trade war, so this model applies to both situations.}
